---
title: "Assignment 5"
author: "Josh martin"
date: "July 12,2016"
output: html_document
---

<style type="text/css">

body, td {
   font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Assignment 5 Solutions

### Problem 3 Assignment 1
The Fibonacci sequence is the following sequence, where the first two terms are both 1 and each term thereafter is the sum of the previous two terms: 1, 1, 2, 3, 5, 8, 13, etc. Write a little R program to find and print the first 30 terms of this sequence. Print them as a simple table that indicates which of these are even, and which are divisible by 3. (Hint: Notice how a number is tested for being even in the code for the Collatz conjecture example.)

#### Solution
```{r}
fib.old <- 0
fib <- 1
fib.new <- 1
for (i in 1:30) {
    cat(i, "\t", fib.new)
    if (fib.new %% 2 == 0)
       cat(" even")
    if (fib.new %% 3 == 0)
       cat(" div3 ")
    cat("\n")
    fib.new <- fib + fib.old
    fib.old <- fib
    fib <- fib.new
}
```



### Problem 1 Assignment 2
Square roots can be found using the following algorithm. Suppose we wish to find the square root of a positive number a. Let the first guess or approximation of the square root be x1 = 1. Then let the next approximation be x2 = (x1 + a/x1)/2. Then let the third approximation be x3 = (x2 + a/x2)/2. If we keep going, the successive approximations will converge quickly to the square root of a.

For example, if we wish to find 31/2, we end up with

x1 = 1
x2 = 2
x3 = 1.75
x4 = 1.732143
x5 = 1.732051 
```{r}
f <- function(a, digits=4) {
  if (a < 0)
    stop("input must be nonnegative\n")
  x <- 1
  old.x <- a
  i <- 1
  tolerance <- 10^(-digits)
  while (abs(x - old.x)>tolerance) {
    old.x <- x
    cat(i,", ", round(x,digits=digits),"\n",sep="")
    x <- (x + a/x)/2
    i <- i + 1
  }
}
```

We verify this does what it's meant to do:
```{r}
f(3)
```

```{r}
f(3,digits=6)
```

```{r}
f(50,digits=3)
```

```{r eval=FALSE}
f(-4)
```
Error in f(-4): input must be nonnegative


### Problem 4 Assignment 2
Create a list that contains the following:

a vector patients containing the names of five patients;
a vector age containing ages in years for these patients;
a vector gender containing codes "M" or "F" for each patient;
a vector BMI containing BMI (body mass index) for each patient (these are typically between 18 and 40); and
a vector TCH containing total cholesterol values for each patient (these are typically between 100 and 300).
(Use your imagination to fill in the data.)

Demonstrate two different ways of accessing: The name of the third patient; the TCH of the fourth patient.
```{r}
patient <- c("Hailey Lewis", "Alejandro Walker", "Jayden Clark", 
           "Zoey Martinez", "Sophia Rogers")
age <- c(46, 32, 20, 68, 42)
gender <- c("F", "M", "M", "F", "F")
bmi <- c(23, 29, 27, 32, 37)
TCH <- c(192, 203, 298, 196, 221)
L <- list(patient=patient, age=age, gender=gender,
          bmi=bmi, TCH=TCH)
```


Now we can access the name of the third patient:
```{r}
L[[1]][3]
```

```{r}
L$patient[3]
```

or the total cholesterol of the fourth patient:

```{r}
L[[5]][4]
```

```{r}
L$TCH[4]
```


### Problem 4 Assignment 4
The coordinates for this problem are given by the following R code:
```{r}
# star 1
x1 <- c(0.954, 4.496, 0.815, 3.23, 3.004, 0.954)
y1 <- c(2.731, 4.119, 5.078, 2.138, 5.935, 2.731)
# star 2
x2 <- c(4.259, 9.934, 4.993, 7.314, 8.5, 4.259)
y2 <- c(4.78, 5.376, 8.229, 3.016, 8.598, 4.78)
```

Produce the following plot. The respective line widths are 6 and 12. (The line widths in the legend are both 5.) The two star colors are <code>"lightsalmon"</code> and <code>"lightgoldenrod"</code> respectively. You may use different colors if you wish, but only if they're interesting (not <code>"red"</code> or <code>"blue"</code>, for example). Google might find you cheat sheets showing R colors.

#### Solution

```{r eval=FALSE, fig.width=5, fig.height=5, echo=TRUE}
plot(x1,y1,col="lightsalmon",type="l",
     lwd=6,xlim=c(0,10),ylim=c(0,10),
     main="Stars", xlab="x coord", ylab="y coord")
lines(x2,y2,col="lightgoldenrod",lwd=12)
legend(0,10,c("star 1","star 2"),lty=c(1,1),lwd=5,
       col=c("lightsalmon","lightgoldenrod"))
```

#### Solution

(Here, I added two lines of code at the beginning, just to make sure where my plot gets saved.)

```{r eval=FALSE, fig.width=5, fig.height=5, echo=TRUE}
setwd("C:/Users/Josh/Documents/Data")
pdf("stars.pdf")
plot(x1,y1,col="lightsalmon",type="l",
     lwd=6,xlim=c(0,10),ylim=c(0,10),
     main="Stars", xlab="x coord", ylab="y coord")
lines(x2,y2,col="lightgoldenrod",lwd=12)
legend(0,10,c("star 1","star 2"),lty=c(1,1),lwd=5,
       col=c("lightsalmon","lightgoldenrod"))
dev.off()
```

## Stars Plot
```{r fig.width=5, fig.height=5, echo=FALSE}
plot(x1,y1,col="lightsalmon",type="l",
     lwd=6,xlim=c(0,10),ylim=c(0,10),
     main="Stars", xlab="x coord", ylab="y coord")
lines(x2,y2,col="lightgoldenrod",lwd=12)
legend(0,10,c("star 1","star 2"),lty=c(1,1),lwd=5,
       col=c("lightsalmon","lightgoldenrod"))
```